Generating Images from Captions\\ with Attention

### 2015-11-08 iclr-paper-new.tex
highlighted changes

\section{Experiments}
\vspace{-0.05in}
\subsection{Microsoft COCO}
\vspace{-0.05in}
Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with at least 5 captions. The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model 
%conditioned on a caption 
very challenging. For consistency with related work on caption generation, we used only the first five captions when training and evaluating our model. 
The images were resized to $32 \times 32$ pixels for consistency with other tiny image datasets \citep{krizhevsky_cifar}. In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more
generated images, visit \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.

@@

\subsubsection{Comparison With Other Models}
\vspace{-0.05in}
Quantitatively evaluating generative models remains a challenging task in itself as each method of evaluation suffers from its own specific drawbacks. Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model. To get a better comparison between performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.

@@

As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods. We rank the images in the test set conditioned on the captions based on the variational lower bound of the log-probabilities 
and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.). Perhaps unsurprisingly, generative models did not perform well on the image retrieval task. To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest 
Euclidean distance in the last fully-connected feature space of a VGG-like model~\citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}. The shortlist of test images used for evaluation can be downloaded from \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}. 
Since there are ``easy'' images for which the model assigns high log-probabilities independent of the query caption, 
we instead look at the ratio of 
the likelihood of the image conditioned on the sentence to the likelihood of the image conditioned on the mean sentence representation in the training set, following the retrieval protocol of~\citep{kiros_captions}.
We also found that the lower bound on the test log-probabilities decreased for sharpened images, and that sharpening considerably hurt the retrieval results. Since sharpening changes the statistics of images, the estimated log-probailities of image pixels is not necessarily a good metric.

>>>>

\section{Experiments}
\vspace{-0.05in}
\subsection{Microsoft COCO}
\vspace{-0.05in}
Microsoft COCO \citep{mscoco} is a very large dataset of \FIXME{$82,783$ images}, each annotated with at least 5 captions. The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model 
%conditioned on a caption 
very challenging. For consistency with related work on caption generation, we used only the first five captions when training and evaluating our model. 
The images were resized to $32 \times 32$ pixels for consistency with other tiny image datasets \citep{krizhevsky_cifar}. In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more
generated images, visit \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.

@@

\subsubsection{Comparison With Other Models}
\vspace{-0.05in}
Quantitatively evaluating generative models remains \FIXME{a challenging task in itself} as each method of evaluation suffers from its own specific drawbacks. Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model. To get a better comparison between performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.

@@

As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods. We rank the images in the test set conditioned on the captions based on the variational lower bound of the log-probabilities 
and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.). Perhaps unsurprisingly, generative models did not perform well on the image retrieval task. To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest 
Euclidean distance in the \FIXME{last fully-connected feature space} of a VGG-like model~\citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}. The shortlist of test images used for evaluation can be downloaded from \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}. 
Since there are ``easy'' images for which the model assigns high log-probabilities independent of the query caption, 
we instead look at the ratio of 
the likelihood of the image conditioned on the sentence to the likelihood of the image conditioned on the mean sentence representation in the training set, following the retrieval protocol of~\citep{kiros_captions}.
We also found that the lower bound on the test log-probabilities decreased for sharpened images, and that sharpening considerably hurt the retrieval results. Since sharpening changes the statistics of images, the estimated log-probailities of image pixels is not necessarily a good metric.

### 2015-11-08 supp.tex
highlighted changes

Each parameter in alignDRAW was initialized by sampling from a Gaussian distribution with 
mean $0$ and standard deviation $0.01$. 
The model was trained using \textit{RMSprop} with an initial learning rate of $0.001$. 
For the Microsoft COCO task, we trained our model for $18$ epochs. The learning rate was 
reduced to $0.0001$ after $11$ epochs.
For the MNIST with Captions task, the model was trained for 150 epochs and the 
learning rate was reduced to $0.0001$ after 110 epochs. During each epoch, randomly created $10000$ training samples were used for training.
The norm of the gradients was clipped at $10$ during training 
to avoid the exploding gradients problem.

>>>>

Each parameter in alignDRAW was initialized by sampling from a Gaussian distribution with 
mean $0$ and standard deviation $0.01$. 
The model was trained using \textit{RMSprop} with an initial learning rate of $0.001$. 
For the Microsoft COCO task, we trained our model for $18$ epochs. The learning rate was 
reduced to $0.0001$ after $11$ epochs.
\FIXME{For the MNIST with Captions task, the model was trained for 150 epochs and the 
learning rate was reduced to $0.0001$ after 110 epochs. During each epoch, randomly created $10,000$ training samples were used for training.}
The norm of the gradients was clipped at $10$ during training 
to avoid the exploding gradients problem.
